---
title: "Computational Modeling - Assignment 2"
author: "Riccardo Fusaroli"
date: "29/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Notes from Riccardo's lecture 3 ###
- priors are always idealized, e.g. flat, made from a limited number of observations
- posteriors are messy
- extract values -> to generate expected outcomes
- each sampling procedure is independent from the previous one (so the new samples don't depend on the previous samples but the actual data)



## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. there is a second part at the bottom for next week.

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)
```{r}
### Riccardo: 3 correct answers out of 6 questions

######################## GRID APPROXIMATION #########################
library(rethinking, brms)

#Define the grid
dens <- 20 #why 20?
p_grid <- seq(from = 0 , to = 1 , length.out = dens)

#Define the prior
prior <- rep( 1 , dens ) # Flat prior: set it to 1, we assume we don't know anything
# prior <- ifelse(p_grid < 0.5 , 0 , 1) # UnreasonablyOptimisticPrior
# prior <- dnorm(p_grid, 0.5, 0.1) # SensiblyCenteredAtChance

#Test the prior (predictive prior check)
dens(rbinom(1e4, 6, runif(1e4, 0, 1))) #everything is equally likeliy know as we have a flat prior -> ofc it's not true #6 because we have 6 question -> for each question run the process 10 000 times -> gives distr given the prior
#if we wanted to check the posterior, we would just have to replace runif thing with posterior(?) what?
      #x axis: number of answers correct (0, 1, 2, ....)
      #y: given the umbe rof smaples we have, whta is the proportion which is with        this value, 
#Compute the likelihood at each value in grid
likelihood <- dbinom( 3 , size = 6 , prob = p_grid ) # 3=number of right answers, 6= number of all questions

#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior

#Standardize the posterior (so it sums to 1)
posterior <- unstd.posterior / sum(unstd.posterior)

#Draw the plot
#make dataframe with the grid values, prior, likelohood and posterior
df <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)
#make plot
ggplot(df, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")
```
Answer: 50% chance Riccardo is better than chance (red line on the graph)
```{r}
############# QUADRATIC APPROXIMATION ################
#we start at random, and we go until in every direction it's 'downhill'
qiccardo_qa <-
  rethinking::map(
    alist(
      w ~ dbinom(6, p),  # binomial likelihood, 6 questions in total
      p ~ dunif(0, 1)    # uniform prior
    ), 
    data = list(w = 3)) #3 correct answers

# display summary of quadratic approximation
precis(qiccardo_qa) #0.5 for mean: that's what we also saw in the graph above

```
2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel. They are both simmetrycal, only Mikkel's distribution is narrower,  so it is more sure or something.
  Answer: Josh is the best, Kistian has too little info, it can't really be used for anything.     Riccardo's and Mikkel's posteriors
2a. Produce plots of the prior, and posterior for each teacher.
```{r}
#Riccardo: 3 correct answers out of 6 questions
#Kristian: 2 correct answers out of 2 questions (then he gets bored)
#Josh: 160 correct answers out of 198 questions (Josh never gets bored)
#Mikkel: 66 correct answers out of 132 questions

############### GRID APPROXIMATION ####################

#making a function that we can apply to all teachers
knowledge <- function(answers_correct, answers_all, answer_name){
  dens <- 100 #so the graphs are more smooth
  p_grid <- seq(from = 0 , to = 1 , length.out = dens)
  prior1 <- rep( 1 , dens )
  dens(rbinom(1e4, answers_all, runif(1e4, 0, 1)))
  likelihood <- dbinom( answers_correct , size = answers_all , prob = p_grid )
  unstd.posterior <- likelihood * prior1
  posterior <- unstd.posterior / sum(unstd.posterior)
  df <- data.frame(grid = p_grid, posterior = posterior, prior = prior1, likelihood = likelihood)
  ggplot(df, aes(grid,posterior)) +  
    geom_point() + 
    geom_line() + 
    theme_classic() + 
    geom_line(aes(grid, prior/dens),color= 'red') + 
    xlab("Knowledge of CogSci") + 
    ylab("posterior probability")+
    ggtitle(answer_name)
  #return(posterior)
}

#Riccardo
knowledge(3,6,"Riccardo")

#Kristian
knowledge(2,2,"Kristian")

#Josh
knowledge(160, 198,"Josh")

#Mikkel
knowledge(66, 132,"Mikkel")
#Riccardo
knowledge(3,6,"Riccardo")

sample(p_grid,pob=posterior, )
quantile(df, 0.5)
```

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.
```{r}
#making a function that we can apply to all teachers
knowledge2 <- function(answers_correct, answers_all, answer_name){
  dens <- 100 #so the graphs are more smooth
  p_grid <- seq(from = 0 , to = 1 , length.out = dens)
  prior2 <- dnorm(p_grid, 0.8, 0.2)
  dens(rbinom(1e4, answers_all, rnorm(1e4, 0, 1)))
  likelihood <- dbinom( answers_correct , size = answers_all , prob = p_grid )
  unstd.posterior <- likelihood * prior2
  posterior <- unstd.posterior / sum(unstd.posterior)
  df <- data.frame(grid = p_grid, posterior = posterior, prior = prior2, likelihood = likelihood)
  ggplot(df, aes(grid,posterior)) +  
    geom_point() + 
    geom_line() + 
    theme_classic() + 
    geom_line(aes(grid, prior/dens),color= 'red') + 
    xlab("Knowledge of CogSci") + 
    ylab("posterior probability")+
    ggtitle(answer_name)
  #return(posterior)
}

#Riccardo
knowledge2(3,6,"Riccardo") #red line is prior and black is the posterior - now our prior is not flat and explains more

#Kristian
knowledge2(2,2,"Kristian")

#Josh
knowledge2(160, 198,"Josh")

#Mikkel
knowledge2(66, 132,"Mikkel") #Mikkel's posterior is more certain (narrower), good thing, it shows that out model has actually used the data and it learned something from it :)
```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?
```{r}
#function 1 (flat prior)
knowledge <- function(answers_correct, answers_all,answer_name){
  dens <- 100 #so the graphs are more smooth
  p_grid <- seq(from = 0 , to = 1 , length.out = dens)
  prior1 <- rep( 1 , dens )
  dens(rbinom(1e4, answers_all, runif(1e4, 0, 1)))
  likelihood <- dbinom( answers_correct , size = answers_all , prob = p_grid )
  unstd.posterior <- likelihood * prior1
  posterior <- unstd.posterior / sum(unstd.posterior)
  df <- data.frame(grid = p_grid, posterior = posterior, prior = prior1, likelihood = likelihood)
  ggplot(df, aes(grid,posterior)) +  
    geom_point() + 
    geom_line() + 
    theme_classic() + 
    geom_line(aes(grid, prior/dens),color= 'red') + 
    xlab("Knowledge of CogSci") + 
    ylab("posterior probability") +
    ggtitle(answer_name)
  #return(posterior)
}
#function 2 - 0.8 mean 0.2 SD
knowledge2 <- function(answers_correct, answers_all, answer_name){
  dens <- 100 #so the graphs are more smooth
  p_grid <- seq(from = 0 , to = 1 , length.out = dens)
  prior2 <- dnorm(p_grid, 0.8, 0.2)
  dens(rbinom(1e4, answers_all, rnorm(1e4, 0, 1)))
  likelihood <- dbinom( answers_correct , size = answers_all , prob = p_grid )
  unstd.posterior <- likelihood * prior2
  posterior <- unstd.posterior / sum(unstd.posterior)
  df <- data.frame(grid = p_grid, posterior = posterior, prior = prior2, likelihood = likelihood)
  ggplot(df, aes(grid,posterior)) +  
    geom_point() + 
    geom_line() + 
    theme_classic() + 
    geom_line(aes(grid, prior/dens),color= 'red') + 
    xlab("Knowledge of CogSci") + 
    ylab("posterior probability") +
    ggtitle(answer_name)
  #return(posterior)
}

### results with functions 1&2
#Riccardo
knowledge(300,600, "Riccardo")
knowledge2(300,600,"Riccardo")

#Kristian
knowledge(200,200,"Kristian")
knowledge2(200,200,"Kristian")

#Josh
knowledge(16000, 19800,"Josh")
knowledge2(16000, 19800,"Josh")

#Mikkel
knowledge(6600, 13200,"Mikkel")
knowledge2(6600, 13200,"Mikkel") #has a more definite peak

### Results in general are not that different
```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
```{r}
#function 3
knowledge3 <- function(answers_correct, answers_all, answers_name){
  dens <- 100 #so the graphs are more smooth
  p_grid <- seq(from = 0 , to = 1 , length.out = dens)
  prior3 <- dnorm(p_grid, 0.2 , 0.2)
  dens(rbinom(1e4, answers_all, rnorm(1e4, 0, 1)))
  likelihood <- dbinom( answers_correct , size = answers_all , prob = p_grid )
  unstd.posterior <- likelihood * prior3
  posterior <- unstd.posterior / sum(unstd.posterior)
  df <- data.frame(grid = p_grid, posterior = posterior, prior = prior3, likelihood = likelihood)
  ggplot(df, aes(grid,posterior)) +  
    geom_point() + 
    geom_line() + 
    theme_classic() + 
    geom_line(aes(grid, prior/dens),color= 'red') + 
    xlab("Knowledge of CogSci") + 
    ylab("posterior probability")+
    ggtitle(answers_name)
  #return(posterior)
}

### results
#Riccardo
knowledge3(3,6,"Riccardo")
knowledge3(300,600,"Riccardo")

#Kristian
knowledge3(2,2,"Kristian")
knowledge3(200,200,"Kristian")

#Josh
knowledge3(160, 198,"Josh")
knowledge2(160, 198,"Josh")

#Mikkel
knowledge3(66, 132,"Mikkel")
knowledge2(66, 132,"Mikkel")
```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?
```{r}

```

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)

d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

# Model sampling only from the prior (for checking the predictions your prior leads to)
FlatModel_priorCheck <- brm(Correct|trials(Questions) ~ 1, 
                 data = subset(d, Teacher=="RF"),
                 prior = prior("uniform(0,1)", class = "Intercept"),
                 family = binomial,
                 sample_prior = "only") # here we tell the model to ignore the data

# Plotting the predictions of the model (prior only) against the actual data
pp_check(FlatModel_priorCheck, nsamples = 100)

# Model sampling by combining prior and likelihood
FlatModel <- brm(Correct|trials(Questions) ~ 1, 
                 data = subset(d, Teacher=="RF"),
                 prior = prior("uniform(0,1)", class = "Intercept"),
                 family = binomial,
                 sample_prior = T)
# Plotting the predictions of the model (prior + likelihood) against the actual data
pp_check(FlatModel, nsamples = 100)

# plotting the posteriors and the sampling process
plot(FlatModel)


PositiveModel_priorCheck <- brm(Correct|trials(Questions) ~ 1,
                     data = subset(d, Teacher=="RF"),
                     prior = prior("normal(0.8,0.2)", 
                                   class = "Intercept"),
                     family=binomial,
                     sample_prior = "only")
pp_check(PositiveModel_priorCheck, nsamples = 100)

PositiveModel <- brm(Correct|trials(Questions) ~ 1,
                     data = subset(d, Teacher=="RF"),
                     prior = prior("normal(0.8,0.2)", 
                                   class = "Intercept"),
                     family=binomial,
                     sample_prior = T)
pp_check(PositiveModel, nsamples = 100)
plot(PositiveModel)

SkepticalModel_priorCheck <- brm(Correct|trials(Questions) ~ 1, 
                      data = subset(d, Teacher=="RF"),
                      prior=prior("normal(0.5,0.01)", class = "Intercept"),
                      family=binomial,
                      sample_prior = "only")
pp_check(SkepticalModel_priorCheck, nsamples = 100)

SkepticalModel <- brm(Correct|trials(Questions) ~ 1, 
                      data = subset(d, Teacher=="RF"),
                      prior = prior("normal(0.5,0.01)", class = "Intercept"),
                      family = binomial,
                      sample_prior = T)
pp_check(SkepticalModel, nsamples = 100)
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)

